---
title: "Keerthi CH Week2 Assignment"
output: pdf_document
date: '2022-07-15'
---

Introduction

For this assignment i have worked on the provided zillow_price.csv dataset. Firstly, installed the necessary libraries and setup our working directory and loading the data and converting the file to a data table, so that it will be easier to look into the data.

```{r libraries and working directory}
# load the data.table, ggolot2, and dplyr libraries and the zillow_price.csv file

library(data.table)
library(dplyr)
library(ggplot2)

setwd("~/Downloads")
dt <- read.csv("zillow_price.csv")

# Convert the file to a data table
dt <- as.data.frame(dt)
```

Methods and Results

In this we are performing some histograms, boxplots, correlation, ggplot, Pearson correlation, linear model

Here we are seeing how many columns we have in our data set and also the summary and structure of the object.

```{r summary}
# how many observations and columns are there?
dim(dt)
ncol(dt)

# use str and summary to see how many missing values we have,
# and what the data looks like
str(dt)
summary(dt)
```

The results shows that we have 90,275 observations and 60 columns in the data table.
And the summary gives the information about the number of Null values in each column.

In this next step we are taking the columns which are numeric and don't have lot of missing values.

```{r numeric cols}
# columns that are numeric and don't have lots of missing values
# you can add others if you like
numeric_cols <- c('bathroomcnt',
                  'bedroomcnt',
                  'calculatedfinishedsquarefeet',
                  'roomcnt',
                  'yearbuilt',
                  'taxvaluedollarcnt',
                  'landtaxvaluedollarcnt',
                  'price')

dt_rows = nrow(dt)                                                              

# Simplify your dataset by only selecting the columns of your choosing dt[, numeric_cols, with = FALSE]

dt <- dt[, numeric_cols]                                         
dt <- na.omit(dt)                                                             
#dt <- dt[complete.cases(dt),]                                                  
dt_rows - nrow(dt)                                                              
```

So, here we see that we have dropped 776 rows which has Null values.

Here we are checking the correlation for home price and the taxvaluedollarcnt and i see that the correlation was found to be 0.95

Created a boxplot and histogram of the price data.

```{r boxplot and histogram}
# We want to try to correlate home price with another variable.
# Create a boxplot of the price data

cor(dt$price, dt$taxvaluedollarcnt)

hist(dt$price)
boxplot(dt$price)

```

Here we are removing the outliers.

```{r outlier}
# Remove the outliers. dt[!which(dt$price %nin% boxplot(dt$price)$out)]

`%nin%` <- Negate(`%in%`)

dt1 <- dt[which(dt$price %nin% boxplot(dt$price)$out),]

```

```{r outliers}

`%nin%` <- Negate(`%in%`)
dt1 <- dt1 %>%
  mutate(price = price / 1000)

dt1 <- dt1 %>%
  mutate(taxvaluedollarcnt = taxvaluedollarcnt / 1000)


boxplot(dt1$price,
        col = "blue")

```

The boxplot shows us that there are houses ranging from very cheap to around $1M.

```{r normalize}
# How many outliers did we drop?
nrow(dt) - nrow(dt1)
```

Here it looks like we have dropped 6020 outliers.

As there are many observations lets take 200 samples and do some plotting.

```{r sample}
# In our case, we have too many observations.  
# Use sample() to only sample a few hundred points to plot.

dt_sample <- dt1[sample(nrow(dt1), size = 200),]
```

Here i have selected the variables price & taxvaluedollarcnt and performed the correlation..

```{r variables}
# plot a few of the more interesting pairs together
cor(dt1$price, dt1$taxvaluedollarcnt)
```

From the below results we can see that the correlation for price and taxvaluedollarcnt was found to be 0.91.

```{r}
# bonus: try to make some nice-looking scatter plots with ggplot2

ggplot(dt_sample, aes(yearbuilt, price)) +
geom_point() +
labs( title = "The relationship between yearbuilt and price")+ geom_smooth()

```

```{r missing values}
# create a new data.table by dropping any missing values
# use dim() to see how many cases we dropped
colSums(dt1==0)

dim(dt1)
```

From the above results we can see that the roomcnt has more number of zeros compared to others.

Here we are performing the pearson correlation for price and taxvaluedollarcnt and we see that the correlation was found to be 0.90 which is a positive correlation.

```{r spearman}
# get the pearson correlation between price and another variable using cor()
#...there are other types of correlations
# try ?cor to see options, and try another correlation 

cor(dt1$price, dt1$taxvaluedollarcnt, method = "spearman")
```

Created a linear model (lm) for a correlated variable by using the price and the taxvaluedollarcnt.
Plotted the variables.

```{r lm}
# use the lm() command to fit a linear model of price to the 
# one variable you think is most correlated or predictive of price
# lm stands for 'linear model'

fit <- lm(dt_sample$taxvaluedollarcnt ~ dt_sample$price)

# view the model summary 
summary(fit)

# plot a scatter plot of the price and the variable you chose

ggplot(dt_sample, aes(taxvaluedollarcnt, price)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method = 'lm', se = FALSE, data = dt_sample)
  labs( title = "Tax value to Price")
```

Added a regression line using abline and below is the plot.

```{r regression line}
# add the regression line to the current plot using abline()
plot(dt_sample$price, dt_sample$taxvaluedollarcnt)
abline(fit, col="red")
```
We can see that they are identical with the exception of the graphics, both indicating a good fit for the model. The ggplot2 model is definitely prettier but with some effort plot can be just as nice.

```{r fit diagnostics}
# plot the fit diagnostics here
par(mfrow=c(2,2)) # Change the panel layout to 2 x 2
plot(fit)

```

The Residual vs fitted plot can show if there is a non-linear relationship between our variables. But here we can see a flat, straight line indicating a linear relationship.

Normal Q-Q plot looks for normal distribution. we can see it is mostly normally distributed, but i think this indicates a slight skewing to the right where the lower priced houses are the exception.

The scale location plot shows the spread of residuals. A flat line would mean a uniform spread. However our data shows a higher spread as the value increases.

The Residuals vs Leverage plot shows extreme value cases can have on the regression line.


Conclusion:

I have downloaded the dataset and converted it to the data table so that it will be easier to look at the data, and i see that there are 90,275 observations and 60 columns in the data table and the summary gave the information about the number of Null values in each column. Next i have dropped 776 rows which has Null values.

Next i have checked the correlation for price and the taxvaluedollarcnt and i see that the correlation was found to be 0.95

Performed box plot to see the outliers in a price column and removed them, and after removing i have plotted the boxplot again and see that the there are houses ranging from very cheap to around $1M approximately.

Observation : I have performed the correlation between price and taxvaluedollarcnt using pearson model and see that my results are 0.90 which is really good..

Next i have selected price:Taxvaluedollarcnt and performed a basic linear model and performed a scatter plot and added a regression line using abline and we see that they are identical with the exception of graphics indicating a good fit for the model and see that we have a strong positive correlation, i also found that outliers do have a negative effect on our model.


